Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-11-17 grove-paper.tex
Add a very rough draft of the cursor section

\newcommand{\Edge}[1]{Edge~#1}
\newcommand{\Vertex}[1]{Vertex~#1}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureEditingNestedParts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{NestedAlice}{L}
\newvertex{NestedAlice}{w}
\newedge{NestedBob}{L}
\newcommand{\figureEditingNestedParts}{
\begin{figure}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
%%%%
\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Editing Different Parts of the Code-c-x) to[star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
\end{tikzpicture}
}
\label{fig:Editing Nested Parts of the Code-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice node],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
\end{tikzpicture}
}
\label{fig:Editing Nested Parts of the Code-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"', >bob edge] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
\end{tikzpicture}
}
\label{fig:Editing Nested Parts of the Code-c}
\label{fig:Editing Nested-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"', >merge edge] -> {
    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge node],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Users Editing Nested Parts of the Code.}
\label{fig:Editing Nested Parts of the Code}
\end{figure}
}

@@

Cursor is either on edge or on position

cursors are advisory

@@

>>>>

\newcommand{\Edge}[1]{Edge~#1}
\newcommand{\Vertex}[1]{Vertex~#1}
\newcommand{\figureCode}[1]{\textbf{\texttt{#1}}\vskip1em}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureEditingNestedParts
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{NestedAlice}{L}
\newvertex{NestedAlice}{w}
\newedge{NestedBob}{L}
\newcommand{\figureEditingNestedParts}{
\begin{figure}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Editing Different Parts of the Code-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Different Parts of the Code-c}}}};
%%%%
\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Editing Different Parts of the Code-c-x) to[star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-a);
\end{tikzpicture}
}
\label{fig:Editing Nested Parts of the Code-a}
\figureCode{\hole{} + w * v}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vNestedAlice" [> "\eNestedAlice"', >alice edge, alice node],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-b) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Editing Different Parts of the Code-c-x) to [out=15,in=165,star] (-2em, 0cm |- fig-Editing Nested Parts of the Code-b);
\end{tikzpicture}
}
\label{fig:Editing Nested Parts of the Code-b}
\figureCode{u * v + \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"', >bob edge] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.25\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Editing Nested Parts of the Code-c) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-a) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
\path [draw,->,merge step] (fig-Editing Nested Parts of the Code-b) to [out=15,in=165] (-2em, 0cm |- fig-Editing Nested Parts of the Code-c);
\end{tikzpicture}
}
\label{fig:Editing Nested Parts of the Code-c}
\label{fig:Editing Nested-c}
\figureCode{w * v + \hole{}}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"', >merge edge] -> {
    "\vNestedAlice" [> "\eNestedAlice"', >merge edge, merge node],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Users Editing Nested Parts of the Code.}
\label{fig:Editing Nested Parts of the Code}
\end{figure}
}

@@

Finally we address the question of cursors.

All edits use edge and vertex identifiers.
Thus cursor positions do not affect the interpretation of graph edits.
Cursors are not considered part of the graph representation,
and need not be communicated to other users.

COMMUNICATION

TODO: use Alice and Bob

However as a convienience to other users, the current cursor position
may be transmitted to other users.
In this role, cursors are merely advisory and we can follow the simple
protocol of the sender marking every cursor announcement with a timestamp
and the receiver displaying the cursor position With the most recently received position being the one displayed.
Note that this cursor position may refer to edges or verticies that
do not yet exist on the receivers machine.
In these cases, we can either not display the other user's cursor (and perhaps
have a visual display flagging this fact), or we can display the most recent
cursor that represents a valid position in the local graph (perhaps
shown in a fadded color to show that this cursor is known to not be up to date).
(Or perhaps, all cursors always fade/decay over time like on radar blips.)

@@

REPRESENTATION

For most editing a cursor is represented by a position within a vertex.
For example, in TODO, Bob's cursor might be at the TODO position of the TODO:vertex.
Representing it this way means that if Bob or some other user deletes TODO:vertex
and replaces it with a different vertex, Bob's cursor doesn't move.
We specify a position within a vertex instead of a vertex because TODO.

However, there is one case where we need a more precise cursor.
That is when a vertex has multiple children in a particular position
and we want the cursor to be on a specific child.
In that case, we represent the cursor by an edge identifier.
An example of this is seen in REF:figure.
If Bob's cursor is on the \texttt{\{ x | y \}} as a whole,
Bob's cursor would be the pair of the vertex id TODO and the position~\texttt{R}.
If Bob's cursor is on a particular element in the multi-child,
say the~\texttt{x}, then we represent the cursor position with edge identifier TODO:edge.
